Exploring Multi-Scale Spatiotemporal Twitter User Mobility Patterns with a Visual-Analytics Approach

### 2016-09-08 MovePattern_IJGI/IJGI_revision/Twitter_MovementPatterns_VisualAnalytics_v2.tex
half

Many recent studies have adopted the LBSM data streams to study human mobility patterns.
For example, they modeled and extracted trajectories of individuals and performed statistical analysis focusing on the distance decay effects in the collective user movements~\cite{gonzalez2008understanding}, which were used to reveal different travel modes~\cite{Jurdak2015}, travel demands~\cite{wu2014intra,hasan2013understanding}, and the impact of social connections~\cite{cho2011friendship}.
These studies have provided strong supports for using LBSM data as proxies for studying mobility patterns of individuals and valuable insights into human mobility dynamics.
However, the variations of movements in different spatial scales and temporal ranges are neglected in these studies, where the measurements of distances are either fixed in a certain time range or to a specific geographic region.
For instances, the examinations on whether there are temporal (e.g., monthly or seasonal) changes within the movements or how the observed mobility patterns vary across different spatial scales (e.g., intra- or inter city or national levels), are lacking.
Such insights are critical to advance our understandings of the collective mobility patterns for various applications, such as examining the mobility patterns across different cities~\cite{noulas2012tale}, the spread patterns of disease~\cite{balcan2009multiscale, tamerius2011global} and touristic activities~\cite{hawelka2014geo}.
On the other hand, while the high-resolution spatiotemporal records from LBSM present unique research opportunities in this direction, the inherited large data volume poses significant data-intensive challenges for developing multi-scale spatiotemporal analysis approaches to dealing with the complexities in filtering movements of individuals, modeling and aggregating user trajectories at multiple spatial and temporal scales~\cite{tsou2015}.

@@

\section{Materials and Methods}
\subsection{Geo-located Twitter data}
Geo-located tweets are tweets appended with an additional geo-tag in the form of a pair of geographical coordinates, which represents the location a tweet was sent at.
In this study, the geo-located tweets were downloaded using the Twitter Streaming API, where we specified a geographical bounding box as an area-of-interest to retrieve all the geo-located tweets that fall within it.
To ensure complete coverage over the continuous United States, we implemented a crawler that selects North America as the initial area-of-interest, where the geographical boundary is specified with lower left (latitude: 5.4, longitude: -167.3) and upper right (latitude: 83.2, longitude: -52.2).
The crawler is constantly running with over 2 million geo-located raw tweets ($\sim$2GB in size) collected per day.
We have collected more than 1.3 billion geo-located tweets from 1$^{st}$ January to 31$^{st}$ December, 2014 with 6,147,430 Twitter users and 1 TB in file size.

@@

To remove non-human users based on unusual relocating speed, a user will be removed if the speed between any two consecutive locations in the user's trajectory with $speed$ $(loc_{i} - loc_{i-1}) > 240 m/s $.
Based on this definition for modeling the space-time Twitter user trajectories, we converted the process of extracting trajectories from the raw geo-located Twitter data as a MapReduce task. Specifically, each mapper utilizes the unique user id as a key to prepares the records that belong to the same user and send them to reducer. Once the reducers receive the $\langle key, value\rangle$ pairs, a Twitter user's space-time trajectory is formed by the sorting the locations in chronological order while considering the speed limit.
\newline

@@

In particular, the measurements of displacements and radius of gyrations of individuals are two popular metrics to investigate and quantify the distance decay effects in the collective mobility patterns~\cite{gonzalez2008understanding}.
The displacement refers to an individual's re-allocation across the geographic space measured in distance, i.e., $distance (loc_{i} - loc_{i-1})$.
It is not equivalent to a ``trip'' took by an individual, for instance, even the time interval between two recorded locations is one month, it will still count as a displacement.
By studying the collective displacements from a group people, it helps to identify the distance bounds associated with different travel modes~\cite{Jurdak2015} and to quantitatively differentiate the mobility patterns from random walks~\cite{brockmann2006scaling}.
On the other hand, radius of gyration (donated as $r_{g}$) is a metric to distinguish mobility patterns of individuals~\cite{gonzalez2008understanding}, which is defined as fellows.
\newline

@@

In particular, as it is shown in Figure~\ref{fig:displacement} (a), the probability distribution of the overall displacements is approximated by $P(d) \sim \lambda_{1} e^{-\lambda_{1}(d - d_{min})}, d_{min}=10m$ from [10 m, 70 m] (accounting for 2 $\%$ of the population), $ P(d) \sim \beta\lambda_{1}d^{\beta-1}e^{-\lambda^{1}(d^\beta-d_{min}^\beta)}, d_{min} = 100m$ from [100 m, 80 km] (accounting for 93 $\%$ of the population), and $P(d) \sim {d}^{-\alpha}$ [$>$ 80 km] (accounting for 5 $\%$ of the population).
In addition, the displacement in the distance bound from 100 m and 80 km in Figure~\ref{fig:displacement} (b) can be further approximately by two power-law distributions with a cutting point at 5 km (53$\%$ distances are less than 5 km and 40$\%$ distances between 5 km and 80 km), which indicates two different travel modes, such as inter- or intra-city movements.
Overall, the fitting functions with different distance bounds suggest the existence of multi-scale or multi-modal mobility patterns~\cite{Jurdak2015} of the Twitter users in the continuous United States, for example the displacements larger than 80 km could be related to inter-state travels or travel by flight.

The probability distribution of radius of gyrations of individuals at the national level (Figure~\ref{fig:gyration} (a)) is approximated by $P(r_{g}) \sim \lambda_{2} e^{-\lambda_{2}(r_{g} - {r_{g}}_{min})}, {r_{g}}_{min}=10m$ from [10 m, 50 m], $P(r_{g}) \sim \lambda_{2} e^{-\lambda_{2}(r_{g} - {r_{g}}_{min})}$ from [50 m, 30 km], and $P(r_{g}) \sim {r_{g}}^{-\alpha}$ [$>$ 30 km].
In particular, the radius of gyration between 50 m and 30 km can be further approximately by two power law distributions with a cutting point at 6 km (Figure~\ref{fig:gyration} (b)), which suggest two main types of spatial coverage of from the collected Twitter users in the continuous United States. 
The distribution shows that around $10\%$ the tweet population has a radius of gyration less than 50 meters, which indicates those twitter users mostly tweet at a particular place, such as home or office;  around 60$\%$ of the population has a radius of gyration less than 30 km, which indicates that most of the collected Twitter user movements are ``short" distances, e.g., within a city locale. Note that the accuracy of these values for defining the distance bound depends on the accuracy of the location information of each geo-located tweet.

@@

\subsection{The interactive 3D virtual globe web mapping interface}
In addition to providing supports for understanding Twitter user mobility patterns with statistical analysis, the framework integrates a 3D virtual globe interface to enable users to perform exploratory geo-visual analytics of the multi-level spatiotemporal Twitter user movements. 
The 3D virtual globe is developed and extended from the Cesium library\footnote{http://cesiumjs.org/}, which is an open-source WebGL virtual globe and map engine. 
We customized the map engine to adapt different spatial scales, which correspond to the hierarchical spatial layers, for aggregating movements in different level-of-details.
The map interface interprets user's interactions, such as area-of-interest, time window, and zoom levels, etc. as parameters and send to the dedicated visualization servlet on the CyberGIS Gateway\footnote{http://sandbox.cigi.illinois.edu/home/}, which is the leading online cyberGIS environment for a large number of users to perform computing- and data-intensive, and collaborative geospatial problem-solving enabled by advanced cyberinfrastructure~\cite{liu2014cybergis}.
In return, the map interface visualizes the corresponding movement flows on the virtual globe.

An overview of the 3D web mapping interface is shown in Figure~\ref{fig:Web_Interface}.
In terms of performing exploratory visual-analytics of Twitter user movement patterns, users can specify the time window to enable the query. 
When the results are shown, users can hover the mouse over each individual lines on the map to see the value of movement flows for both in and out directions.
If the selected criteria keep unchanged, whenever the user zooms in/out, tilt or rotate the globe, the 3D virtual globe mapping interface will automatically provide the corresponding level-of-details on the fly.
For example, Figure~\ref{fig:chicago} and Figure~\ref{fig:Chicago_airport} demonstrated the movement flows in different level-of-details around the Chicago city, and between O'Hare International Airport and the city center of Chicago city, where top 20 $\%$ movement flows were shown.
The URL to access this 3D web mapping interfaces as well as the source code of the visual-analytics framework are available upon request.

@@

We studied the distance decay effects in the collective Twitter user movements measured by the probability distributions of the displacements and radius of gyrations of individuals. 
These distributions can all be approximated by a combination of three functions: an exponential function, a stretched-exponential function and a power-law function. In particular, distance bounds between different fitting functions in displacement distribution reveals the existence of multi-scale or multi-modal mobility patterns of the Twitter users, whereas the distribution of radius of gyration reveals different groups of Tweet users with different types of spatial coverages at multiple spatial scales.   
We further studied these mobility patterns in different temporal ranges to investigate the temporal changes in the mobility patterns. 
We found that the accumulated displacements are almost identical in different months, while the long distance radius of gyration (i.e., above 30 km) will experience more seasonal changes in the Twitter user mobility pattern.

>>>>

Many recent studies have adopted the LBSM data streams to study human mobility patterns.
For example, they modeled and extracted trajectories of individuals and performed statistical analysis focusing on the distance decay effects in the collective user movements~\cite{gonzalez2008understanding}, which were used to reveal different travel modes~\cite{Jurdak2015}, travel demands~\cite{wu2014intra,hasan2013understanding}, and the impact of social connections~\cite{cho2011friendship}.
These studies have provided strong supports for using LBSM data as proxies for studying mobility patterns of individuals and valuable insights into human mobility dynamics.
However, the variations of movements in different spatial scales and temporal ranges are neglected in these studies, where the measurements of distances are either fixed in a certain time range or to a specific geographic region.
For instances, the examinations on whether there are temporal (e.g., monthly or seasonal) changes within the movements or how the observed mobility patterns vary across different spatial scales (e.g., intra- or inter city or national levels) are lacking.
Such insights are critical to advance our understandings of the collective mobility patterns for various applications, such as examining the mobility patterns across different cities~\cite{noulas2012tale}, the spread patterns of disease~\cite{balcan2009multiscale, tamerius2011global} and touristic activities~\cite{hawelka2014geo}.
On the other hand, while the high-resolution spatiotemporal records from LBSM present unique research opportunities in this direction, the inherited large data volume poses significant data-intensive challenges for developing multi-scale spatiotemporal analysis approaches to dealing with the complexities in filtering movements of individuals, modeling and aggregating user trajectories at multiple spatial and temporal scales~\cite{tsou2015}.

@@

\section{Materials and Methods}
\subsection{Geo-located Twitter data}
Geo-located tweets are tweets appended with an additional geo-tag in the form of a pair of geographical coordinates, which represents the location a tweet was sent at.
In this study, the geo-located tweets were downloaded using the Twitter Streaming API, where we specified a geographical bounding box as an area-of-interest to retrieve all the geo-located tweets that fall within it.
To ensure complete coverage over the continuous United States, we implemented a crawler that selects North America as the initial area-of-interest, where the geographical boundary is specified with lower left (latitude: 5.4, longitude: -167.3) and upper right (latitude: 83.2, longitude: -52.2).
The crawler is constantly running with over 2 million geo-located raw tweets ($\sim$2 GB in size) collected per day.
We have collected more than 1.3 billion geo-located tweets from 1$^{st}$ January to 31$^{st}$ December, 2014 with 6,147,430 Twitter users and 1 TB in file size.

@@

To remove non-human users based on unusual relocating speed, a user will be removed if the speed between any two consecutive locations in the user's trajectory with $speed$ $(loc_{i} - loc_{i-1}) > 240 m/s $.
Based on this definition for modeling the space-time Twitter user trajectories, we converted the process of extracting trajectories from the raw geo-located Twitter data as a MapReduce task. Specifically, each mapper utilizes the unique user id as a key to prepares the records that belong to the same user and send them to a reducer. Once the reducers receive the $\langle key, value\rangle$ pairs, a Twitter user's space-time trajectory is formed by the sorting the locations in chronological order while considering the speed limit.
\newline

@@

In particular, the measurements of displacements and radius of gyrations of individuals are two popular metrics to investigate and quantify the distance decay effects in the collective mobility patterns~\cite{gonzalez2008understanding}.
The displacement refers to an individual's re-allocation across the geographic space measured in distance, i.e., $distance (loc_{i} - loc_{i-1})$.
It is not equivalent to a ``trip'' took by an individual, for instance, even the time interval between two recorded locations is one month, it will still count as a displacement.
By studying the collective displacements from a group people, it helps to identify the distance bounds associated with different travel modes~\cite{Jurdak2015} and to quantitatively differentiate the mobility patterns from random walks~\cite{brockmann2006scaling}.
On the other hand, radius of gyration (donated as $r_{g}$) is a metric to distinguish mobility patterns of individuals~\cite{gonzalez2008understanding}, which is defined as follows.
\newline

@@

In particular, as it is shown in Figure~\ref{fig:displacement} (a), the probability distribution of the overall displacements is approximated by $P(d) \sim \lambda_{1} e^{-\lambda_{1}(d - d_{min})}, d_{min}=10~m$ from [10 m, 70 m] (accounting for 2 $\%$ of the population), $ P(d) \sim \beta\lambda_{1}d^{\beta-1}e^{-\lambda^{1}(d^\beta-d_{min}^\beta)}, d_{min}=100~m$ from [100 m, 80 km] (accounting for 93 $\%$ of the population), and $P(d) \sim {d}^{-\alpha}$ [$>$ 80 km] (accounting for 5 $\%$ of the population).
In addition, the displacement in the distance bound from 100 m and 80 km in Figure~\ref{fig:displacement} (b) can be further approximately by two power-law distributions with a cutting point at 5 km (53$\%$ distances are less than 5 km and 40$\%$ distances between 5 km and 80 km), which indicates two different travel modes, such as inter- or intra-city movements.
Overall, the fitting functions with different distance bounds suggest the existence of multi-scale or multi-modal mobility patterns~\cite{Jurdak2015} of the Twitter users in the continuous United States, for example the displacements larger than 80 km could be related to inter-state travels or travel by flight.

The probability distribution of radius of gyrations of individuals at the national level (Figure~\ref{fig:gyration} (a)) is approximated by $P(r_{g}) \sim \lambda_{2} e^{-\lambda_{2}(r_{g} - {r_{g}}_{min})}, {r_{g}}_{min}=10~m$ from [10 m, 50 m], $P(r_{g}) \sim \lambda_{2} e^{-\lambda_{2}(r_{g} - {r_{g}}_{min})}$ from [50 m, 30 km], and $P(r_{g}) \sim {r_{g}}^{-\alpha}$ [$>$ 30 km].
In particular, the radius of gyration between 50 m and 30 km can be further approximately by two power law distributions with a cutting point at 6 km (Figure~\ref{fig:gyration} (b)), which suggest two main types of spatial coverage of from the collected Twitter users in the continuous United States.
The distribution shows that around $10\%$ the tweet population has a radius of gyration less than 50 meters, which indicates those twitter users mostly tweet at a particular place, such as home or office;  around 60$\%$ of the population has a radius of gyration less than 30 km, which indicates that most of the collected Twitter user movements are ``short" distances, e.g., within a city locale. Note that the accuracy of these values for defining the distance bound depends on the accuracy of the location information of each geo-located tweet.

@@

\subsection{The interactive 3D virtual globe web mapping interface}
In addition to providing supports for understanding Twitter user mobility patterns with statistical analysis, the framework integrates a 3D virtual globe interface to enable users to perform exploratory geo-visual analytics of the multi-level spatiotemporal Twitter user movements\footnote{http://sandbox.cigi.illinois.edu/home/apps.php?app=movepattern}. 
The 3D virtual globe is developed and extended from the Cesium library\footnote{http://cesiumjs.org/}, which is an open-source WebGL virtual globe and map engine. 
We customized the map engine to adapt different spatial scales, which correspond to the hierarchical spatial layers, for aggregating movements in different level-of-details.
The map interface interprets user's interactions, such as area-of-interest, time window, and zoom levels, etc. as parameters and send to the dedicated visualization servlet on the CyberGIS Gateway, which is the leading online cyberGIS environment for a large number of users to perform computing- and data-intensive, and collaborative geospatial problem-solving enabled by advanced cyberinfrastructure~\cite{liu2014cybergis}.
In return, the map interface visualizes the corresponding movement flows on the virtual globe.

An overview of the 3D web mapping interface is shown in Figure~\ref{fig:Web_Interface}.
In terms of performing exploratory visual-analytics of Twitter user movement patterns, users can specify the time window to enable the query. 
When the results are shown, users can hover the mouse over each individual lines on the map to see the value of movement flows for both in and out directions.
If the selected criteria keep unchanged, whenever the user zooms in/out, tilt or rotate the globe, the 3D virtual globe mapping interface will automatically provide the corresponding level-of-details on the fly.
For example, Figure~\ref{fig:chicago} and Figure~\ref{fig:Chicago_airport} demonstrated the movement flows in different level-of-details around the Chicago city, and between O'Hare International Airport and the city center of Chicago city, where top 20 $\%$ movement flows were shown.
The source code of the visual-analytics framework is available upon request.

@@

We studied the distance decay effects in the collective Twitter user movements measured by the probability distributions of the displacements and radius of gyrations of individuals. 
These distributions can all be approximated by a combination of three functions: an exponential function, a stretched-exponential function and a power-law function. In particular, distance bounds between different fitting functions in displacement distribution reveals the existence of multi-scale or multi-modal mobility patterns of the Twitter users, whereas the distribution of radius of gyration reveals different groups of Twitter users with different types of spatial coverages at multiple spatial scales.   
We further studied these mobility patterns in different temporal ranges to investigate the temporal changes in the mobility patterns. 
We found that the accumulated displacements are almost identical in different months, while the long distance radius of gyration (i.e., above 30 km) will experience more seasonal changes in the Twitter user mobility pattern.
